---
title: "DAPT 622 Assignment 1"
author: "Daniel Erb"
date: "January 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load Libraries and Arrange Data}
library(knitr)
library(kableExtra)
library(magrittr)
library(DT)
library(dplyr)
library(ggplot2)
library(GGally)
candy_data <- read.table("G:/My Drive/vcu/Spring 2019/Statistics/Assignments/Assignment 1/DAPT_622_Assignment_1/Assignment_Materials/CandyBars.txt",sep=',',header=TRUE)

cd_sub <- candy_data[,c("Oz.pkg","Calories","Total.fat.g"
                  ,"Saturated.fat.g","Cholesterol.g"
                  ,"Sodium.mg","Carbohydrate.g"
                  ,"Dietary.fiber.g","Sugars.g"
                  ,"Protein.g","Vitamin.A..RDI"
                  ,"Vitamin.C..RDI","Calcium..RDI"
                  ,"Iron..RDI")]

```

# Part 1 - Exploratory Data Analysis

## Question 1

Perform the following exploratory data analysis tasks: 

1. Consider all of the numeric variables (i.e. all of the variables except Brand and Name): Determine the variance/covariance matrix and the correlation matrix of these variables.  Interpret briefly what you see.

```{r Generate Matrices}

#Get variance-covariance matrix
covmat <- cov(cd_sub)
kable(covmat,caption = "Variance/Covariance Matrix for Candy Data", digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = TRUE, font_size = 10)

#Get correlation matrix
corrmat <- cor(cd_sub)
kable(corrmat,caption = "Correlation Matrix for Candy Data", digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = TRUE, font_size = 10)

```
***
The variance/covariance matrix and correlation matrix show several positive correlations of note. For example, higher calorie counts tends to be associated with higher levels of total fat, saturated fat, carbohydrates, dietary fiber, sugars, and protein. Total fat, saturated fat and cholesterol are each positively correlated with one another. Protein correlates positively with total fat as well as dietary fiber, but interestingly is not very correlated with saturated fat or cholesterol. Vitamin A, Vitamin C, Calcium, and Iron are all positively correlated to one another to some extent.

There are very few negative correlations within the candy bar data set. Those negative correlations that do exist are of less magnitude than many of the positive correlations. The strongest negative correlation (-0.305) exists between carbohydrates and cholesterol. Other negative correlations are below -0.25 but do tend to be between vitamins and either carbohydrates, sugars, total fat, saturated fat, or calories. 

```{r Generate Scatterplot Matrix}

ggcorr(cd_sub, palette = "RdBu", label = TRUE, hjust = 0.75)

ggpairs(cd_sub)

ggplot(cd_sub) + 
  plotmatrix()

```
   
*Thanks!*
