---
title: "DAPT 622 Assignment 1"
author: "Daniel Erb"
date: "January 14, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

```{r Load Libraries and Arrange Data, message = FALSE}
library(kableExtra)
library(corrplot)
library(stats)
candy_data <- read.table("G:/My Drive/vcu/Spring 2019/Statistics/Assignments/Assignment 1/DAPT_622_Assignment_1/Assignment_Materials/CandyBars.txt",sep=',',header=TRUE)

cd_sub <- candy_data[,c("Oz.pkg","Calories","Total.fat.g"
                  ,"Saturated.fat.g","Cholesterol.g"
                  ,"Sodium.mg","Carbohydrate.g"
                  ,"Dietary.fiber.g","Sugars.g"
                  ,"Protein.g","Vitamin.A..RDI"
                  ,"Vitamin.C..RDI","Calcium..RDI"
                  ,"Iron..RDI")]

```

# Part 1 - Exploratory Data Analysis

Perform the following exploratory data analysis tasks:

## Question 1

1. Consider all of the numeric variables (i.e. all of the variables except Brand and Name): Determine the variance/covariance matrix and the correlation matrix of these variables.  Interpret briefly what you see.

```{r Generate Matrices, message = FALSE}

#Get variance-covariance matrix
covmat <- cov(cd_sub)
kable(covmat,caption = "Variance/Covariance Matrix for Candy Data", digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = TRUE, font_size = 10)

#Get correlation matrix
corrmat <- cor(cd_sub)
kable(corrmat,caption = "Correlation Matrix for Candy Data", digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = TRUE, font_size = 10)

```
***
The variance/covariance matrix and correlation matrix show several positive correlations of note. For example, higher calorie counts tends to be associated with higher levels of total fat, saturated fat, carbohydrates, dietary fiber, sugars, and protein. Total fat, saturated fat and cholesterol are each positively correlated with one another. Protein correlates positively with total fat as well as dietary fiber, but interestingly is not very correlated with saturated fat or cholesterol. Vitamin A, Vitamin C, Calcium, and Iron are all positively correlated to one another to some extent.

There are very few negative correlations within the candy bar data set. Those negative correlations that do exist are of less magnitude than many of the positive correlations. The strongest negative correlation (-0.305) exists between carbohydrates and cholesterol. Other negative correlations are below -0.25 but do tend to be between vitamins and either carbohydrates, sugars, total fat, saturated fat, or calories. 

## Question 2

2. Construct a scatterplot matrix of the variables specified in #1. Comment on what you observe. 

```{r Generate Scatterplot Matrix, message = FALSE}

pairs(~Oz.pkg
      +Calories
      +Total.fat.g
      +Saturated.fat.g
      +Cholesterol.g
      +Sodium.mg
      +Carbohydrate.g
      +Dietary.fiber.g
      +Sugars.g
      +Protein.g
      +Vitamin.A..RDI
      +Vitamin.C..RDI
      +Calcium..RDI
      +Iron..RDI
      ,data = cd_sub
      #,lower.panel=panel.smooth, upper.panel=panel.cor
      #,pch=20
      , gap = .2
      , main="Candy Scatterplot Matrix")


```

The scatterplot matrix for the candy bar data shows several noteworthy patterens. Ounces could potentially have a linear relationship with carbohydrates. Calories seems to have a linear relationship with total fat. Total fat also seems to be linearly related to saturated fat and protein. Carbohydrates also seem linearly related to sugar. Other relationships may be difficult to assess from a scatterplot with so many attributes represented.

## Question 3

3. Construct a color map on correlations for the variables specified in #1. Comment on what you observe.

```{r Color Map Creation} 

corrmat <- cor(cd_sub)
corrplot(corrmat,method="color")

```

The color map of the correlations show very similar results as the variance/covariance and correlation matrices. They are quite a few strongly positive correlations. There are also far fewer strong negative correlations.

## Question 4

4. Use a probability plot to assess the multivariate normality of the variables specified in #1.  Interpret.

```{r Probability Plot Creation}

x <- as.matrix(cd_sub)
center <- colMeans(x)
n <- nrow(x)
p <- ncol(x)
cov <- cov(x)
d <- mahalanobis(x,center,cov)
qqplot(qchisq(ppoints(n),df = p),d,
       main = "QQ Plot Assessing Multivariate Normality",
       ylab = "Mahalanobis D2",
       xlab = "Expected value")
abline(a = 0, b = 1)
```

The multivariate probability plot shows the Mahalanobis distances aren't too far from the expected values below ~20. Above this point there are 5 candy bars that fall slightly outside the expected distances and another 2 that are well outside of the expected values. These 2 points would most likely need evaluation for their effects on any analysis performed.

---

# Part 2 - Principal Component Analysis

Reconsider the candy bar data. Perform a Principal Component Analysis on this data by carrying out the following steps:

## Section A

### Subsection i

Find and display the eigenvalues of the correlation matrix. Use these along with a Scree plot (and/or other means) to determine the number of principal components that are to be retained. How much of the variation in the data is explained by your chosen number of principal components? Remember that your goal is dimension reduction…so please be sensible in your choice of number of components

```{r Find eigenvalues and scree plot creation}

#Compute the correlation matrix

 ##recall corrmat from above

#compute the eigenvalues and eigenvectors

ev <- eigen(corrmat)
ev$values
ev$vectors

# scale = TRUE allows for correlation matrix to be utilized
# FALSE would allow for covariance matrix to be utilized
pca <- prcomp(cd_sub, scale = TRUE)
summary(pca)

#(show plots)

screeplot(pca, npcs = 14, type = "lines")


#Biplot
biplot(pca)

```


Considering the generated Scree plot as well as the biplot for the eigenvectors of the 14 principal components, 3 principal components should be kept. The first 3 principal components explain nearly 60% (0.5969) of the variability in the data, while keeping the number features relatively low.

### Subsection ii

Provide the loadings matrix.  What do you learn from this matrix? Using this matrix, and as best as possible, interpret the first two principal components (I will expect some interpretation).

```{r loadings for pca}

#Loadings
loadings <- pca$rotation
loadings

```

Principal component 1 shows several major contributors. Calories, total fat, saturated fat, dietary fiber, and protein are the major positive contributors to this component while there are no seemingly significant negative contributors. This would seem to suggests, as we might suspect from previous candy intake, that a candy bars content of calories, total fat, saturated fat, dietary fiber, and protein are positively related in some way.

Principal component 2 shows only a few slightly positive values for carbohydrates and sugars. This is in conjunction with some stronger negative loadings for vitamin A, vitamin C, calcium, and iron. This would suggest that a candy bars carbohydrate and sugar contents are opposed to increased values of vitamin A, vitamin C, calcium and iron.





















---
*Thanks!*
