---
title: "DAPT 622 Assignment 1"
author: "Daniel Erb"
date: "January 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load Libraries and Arrange Data}
library(knitr)
library(kableExtra)
library(magrittr)
library(DT)
library(dplyr)
candy_data <- read.table("G:/My Drive/vcu/Spring 2019/Statistics/Assignments/Assignment 1/DAPT_622_Assignment_1/Assignment_Materials/CandyBars.txt",sep=',',header=TRUE)

cd_sub <- candy_data[,c("Oz.pkg","Calories","Total.fat.g"
                  ,"Saturated.fat.g","Cholesterol.g"
                  ,"Sodium.mg","Carbohydrate.g"
                  ,"Dietary.fiber.g","Sugars.g"
                  ,"Protein.g","Vitamin.A..RDI"
                  ,"Vitamin.C..RDI","Calcium..RDI"
                  ,"Iron..RDI")]

```

# Part 1 - Exploratory Data Analysis

## Question 1

Perform the following exploratory data analysis tasks: 

1. Consider all of the numeric variables (i.e. all of the variables except Brand and Name): Determine the variance/covariance matrix and the correlation matrix of these variables.  Interpret briefly what you see.

```{r Generate Matrices}

#Get variance-covariance matrix
covmat <- cov(cd_sub)
kable(covmat,caption = "Variance/Covariance Matrix for Candy Data", digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = TRUE, font_size = 10)

#Get correlation matrix
corrmat <- cor(cd_sub)
kable(corrmat,caption = "Correlation Matrix for Candy Data", digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = TRUE, font_size = 10)

data.frame(corrmat[,]) %>%
  mutate_if(is.numeric, function(x) {
    cell_spec(x, 
              color = spec_color(x, begin = 0, end = 0.7))
  }) %>%
  kable(escape = F,align = "c") %>%
  kable_styling(c("striped", "condensed"), full_width = F)
```
***
The variance/covariance matrix and correlation matrix show several positive correlations of note. Higher calorie counts tends to be associated with higher levels of total fat, saturated fat, carbohydrates, dietary fiber, sugars, and protein.
   
*Thanks!*
